Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-09-08 paper/main.tex
[ paper ] minor changes, title and abstract.
[ paper ] fixing some mistakes.
[ paper ] fixing error grammars and adding notes.
[ paper ] section 3 introduction added.
[ paper ] appendix added.

\documentclass[runningheads,a4paper]{llncs}

\usepackage{subfiles}

@@

% --------------------------------------------------------------------------
% TODO notes
% --------------------------------------------------------------------------

@@

% --------------------------------------------------------------------------
% Tikz Configuration
% --------------------------------------------------------------------------

@@

% --------------------------------------------------------------------------
% Listings
% --------------------------------------------------------------------------

@@

\lstdefinestyle{TPTP}{
    aboveskip=10pt
  , basicstyle=\tt
  , belowskip=10pt
  , keywords=[2]{axiom, conjecture,inference, negated_conjecture, plain, source}
  , keywords=[3]{fof, cnf}
  , keywords=[4]{canonicalize, resolve, conjunct,
    strip, simplify, negate, clausify}
  , keywordstyle=[2]\bfseries
  , keywordstyle=[3]\bfseries
  , keywordstyle=[4]\bfseries
  , showstringspaces=false
  , keepspaces=true
}

@@

%\usepackage[utf8]{inputenc}
\usepackage{ucs}
\usepackage{autofe}
\usepackage{afterpage}
\usepackage{comment}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{stmaryrd} %new symbol font for tcs
\usepackage{color}
\usepackage{fancybox}

@@

% --------------------------------------------------------------------------
% New commands
% --------------------------------------------------------------------------
\usepackage{xspace}

@@

\newcommand{\prg}[1]{\texttt{#1}\xspace}
\newcommand{\Agda}{\prg{Agda}}
\newcommand{\Metis}{\prg{Metis}}
\newcommand{\IsabelleHOL}{\prg{IsabelleHOL}}

@@

% --------------------------------------------------------------------------
% Rest.
% --------------------------------------------------------------------------

@@

\institute{Universidad EAFIT, Logic and Computation Research Group,\\
Department of Mathematics, Medell\'in, Colombia\\
\email{\{jprieto9,asr\}@eafit.edu.co}}

@@

\subfile{sections/abstract.tex}
\subfile{sections/introduction.tex}
\subfile{sections/metis-language-and-proofs.tex}
% \subfile{sections/proof-reconstruction.tex}
\subfile{sections/related-work.tex}
\subfile{sections/conclusions.tex}
\subfile{sections/acknowledgements.tex}

@@

>>>>

\documentclass[runningheads, a4paper]{llncs}
\usepackage{subfiles}

@@

\usepackage[title]{appendix}

% -------------------------------------------------------------------
% TODO notes
% -------------------------------------------------------------------

@@

% -------------------------------------------------------------------
% Tikz Configuration
% -------------------------------------------------------------------

@@

% -------------------------------------------------------------------
% Listings
% -------------------------------------------------------------------

@@

\lstdefinestyle{TPTP}{
    aboveskip=10pt
  , basicstyle=\tt
  , belowskip=10pt
  , keywords=[2]{axiom
    , conjecture
    , inference
    , negated_conjecture
    , plain
    , source}
  , keywords=[3]{fof, cnf}
  , keywords=[4]{canonicalize, resolve, conjunct,
    strip, simplify, negate, clausify}
  , keywordstyle=[2]\bfseries
  , keywordstyle=[3]\bfseries
  , keywordstyle=[4]\bfseries
  , showstringspaces=false
  , keepspaces=true
}

@@

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{ucs}
\usepackage{autofe}
\usepackage{afterpage}
\usepackage{comment}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{stmaryrd} %new symbol font for tcs
\usepackage{color}
\usepackage{fancybox}

\usepackage{textcomp}
\usepackage[verbose]{newunicodechar}
\newunicodechar{∧}{\ensuremath{\wedge}}
\newunicodechar{∨}{\ensuremath{\vee}}
\newunicodechar{¬}{\ensuremath{\neg}}
\newunicodechar{⇒}{\ensuremath{\Rightarrow}}
\newunicodechar{⇔}{\ensuremath{\Leftrightarrow}}
\newunicodechar{∼}{\ensuremath{\sim}}
\newunicodechar{₀}{\ensuremath{{}_{0}}}
\newunicodechar{₁}{\ensuremath{{}_{1}}}
\newunicodechar{₃}{\ensuremath{{}_{2}}}
\newunicodechar{₄}{\ensuremath{{}_{3}}}
\newunicodechar{₅}{\ensuremath{{}_{4}}}
\newunicodechar{₆}{\ensuremath{{}_{5}}}
\newunicodechar{₇}{\ensuremath{{}_{6}}}
\newunicodechar{₈}{\ensuremath{{}_{7}}}
\newunicodechar{₉}{\ensuremath{{}_{8}}}
\newunicodechar{∅}{\ensuremath{\emptyset}}
\newunicodechar{⊢}{\ensuremath{\vdash}}
\newunicodechar{⊥}{\ensuremath{\bot}}
\newunicodechar{Γ}{\ensuremath{\Gamma}}

%\DeclareUnicodeCharacter{00A0}{~}

@@

% -------------------------------------------------------------------
% New commands
% -------------------------------------------------------------------
\usepackage{xspace}

@@

\newcommand{\prg}[1]{\texttt{#1}\xspace}
\newcommand{\Agda}{\prg{Agda}}
\newcommand{\Metis}{\prg{Metis}}
\newcommand{\Athena}{\prg{Athena}}
\newcommand{\IsabelleHOL}{\prg{IsabelleHOL}}

@@

% -------------------------------------------------------------------
% Rest.
% -------------------------------------------------------------------

@@

\institute{Logic and Computation Research Group,\\
Universidad EAFIT, Medell\'in, Colombia\\
\email{\{jprieto9,asr\}@eafit.edu.co}}

@@

\subfile{sections/abstract.tex}
\subfile{sections/introduction.tex}
\subfile{sections/metis-language-and-proofs.tex}
\subfile{sections/proof-reconstruction.tex}
\subfile{sections/conclusions.tex}
\subfile{sections/related-work.tex}
\subfile{sections/acknowledgements.tex}

@@

\newpage
\subfile{sections/appendices.tex}

### 2017-09-08 paper/sections/conclusions.tex
[ paper ] appendix added.

In this document, we attempted to reconstruct proofs
of propositional logic without equality deliver by \Metis successfully.

From our experience building a proof reconstruction tool, depending 
on the systems involved in the translation, some difficulties can 
arise. We listed some of them in section
\ref{sec:metis-language-and-proofs} and in the section
\ref{sec:proof-reconstruction}.

@@

However, one of our biggest obstacles
\improvement{Remove the use of the \emph{biggest obstacles}.}
was dealing with
the Equality between formulas since, in our approach, we only work
with the syntactic treatment of the logic without considering the
semantics.
One of the main reasons for such a restriction relay on our plans to
support further first-order logic, where a semantic treatment is not
viable.\unsure{Please confirm this.}

>>>>

In this document, we reconstruct proofs
of propositional logic without equality deliver by \Metis 
successfully.

Despite the fact, our logic domain is more straightforward than 
first-order or high-order logic where there are many 
proof-reconstruction tools (see a list of some of them in the 
related work, section \ref{sec:related-work}), we found some issues 
around the reconstruction that show this task is far from trivial.
Depending  on the systems involved in the translation, some 
difficulties can arise. We listed some of them in section
\ref{sec:metis-language-and-proofs} and in the section
\ref{sec:proof-reconstruction}.

@@

An important matter was how dealing with
the equality between formulas since, in our approach, we only work
with the syntactic treatment of the logic without considering the
semantics.
One of the main reasons for such a restriction relay on our plans 
further to support first-order logic, where a semantic treatment is 
not viable since it computational complexity.\unsure{Please confirm this.}

### 2017-09-08 paper/sections/proof-reconstruction.tex
[ paper ] appendix added.

With such a derivation, we process it with our \Haskell translator 
tool, \Athena. \Athena parses the \TSTP format, analyze the 
derivation and generates a representation of the natural deduction 
proof using a version of a tree data structure (see the properties 
of this tree in subsection \ref{ssec:metis-proofs}). As a result,
\Athena generates an \Agda file with such a proof. We have included 
in this file, imports for \Agda libraries that give the support for 
the logic~\cite{AgdaProp}, and versions of the \Metis' inference 
rules~\cite{AgdaMetis}.

Finally, we use the proof assistant in step No. 7 to type-check the 
proof and verify its correctness. We have only one answer from two 
possible answers: the type-checking success (i.e., the proof is 
valid) or not. If the type-check fails, one of the three actors are 
responsible: \Metis, \Athena, or \Agda. We are responsible for 
errors committed in the translation from \TSTP to \Agda; lack of the 
coverage of the \Metis' inference rules, or bad assumptions about 
them.

>>>>

With such a derivation, we process it with our \Haskell translator 
tool, \Athena~\cite{Athena}.
\Athena parses the \TSTP format, analyze the 
derivation and generates a representation of the natural deduction 
proof using a version of a tree data structure (see the properties 
of this tree in subsection \ref{ssec:metis-proofs}). As a result,
\Athena generates an \Agda file with such a proof. We have included 
in this file, imports for \Agda libraries that give the support for 
the logic~\cite{AgdaProp}, and versions of the \Metis' inference 
rules~\cite{AgdaMetis}.

Finally, we use the proof assistant in step No. 7 to type-check the 
proof and verify its correctness. We have only one answer from two 
possible answers: type-checking succeed (i.e., the proof is 
valid) or type-checking failed.
If the type-check fails, one of the three actors are 
responsible: \Metis, \Athena, or \Agda. We are responsible for 
errors committed in the translation from \TSTP to \Agda; lack of the 
coverage of the \Metis' inference rules, or bad assumptions about 
them.
